Hydrogen Epoch of Reionization Array (HERA)

### 2016-04-20 manuscript.tex
Tweaks and polishing of Introduction

The Hydrogen Epoch of Reionization Array (HERA; http://reionization.org) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
other probes, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of a new understanding of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).

Because of these key science questions that may be probed by highly-redshifted hydrogen emission, there has been a resurgence in low frequency arrays  -- most with a primary
objective of measuring the EOR.  These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
With the first generation 21\,cm telescopes approaching the limits of their sensitivity however, HERA continues the experiment by bringing significantly more sensitivity to bear
while continuing to optimize its design along the delay-spectrum approach to detect the power spectrum of the EoR (discussed in \S\ref{sec:delayapproach}).
At the same time, HERA continues the development of new techniques to extend and improve its capabilities.
Examples include improving foreground subtraction to expand the information content 
in power spectral and imaging measurements and
extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).

@@

The HERA experiment will comprise 
350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
Astronomy Reserve (see Fig. \ref{fig:HERApictures} (top) for a rendering).
HERA has been funded under the US National Science Foundation's Mid-Scale Innovations Program to begin building elements optimized for robust power spectrum detection in a 
staged deployment.  The key is to produce inexpensive collecting area optimized to detect the EOR.  
The first 19-element array is complete and, with additional funding from Cambridge University, construction underway on the next 18 elements, for a total of 37.
The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in the construction and science.

The current 19 elements are testing the HERA element design for performance and manufacturability on location.  The 37 elements, to be completed in 2016, will provide a factor of about 5 more 
sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
%\citep{ali_et_al2015}
which should provide a very robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.  These estimates will be quantified and shown in this paper.

HERA is a collaboration of the following Partner institutions:  
Arizona State University (Tempe, AZ USA), 
Brown University (Providence, RI USA), 
University of California Berkeley (Berkeley, CA USA), 
University of California Los Angeles (Los Angeles, CA USA), 
University of Cambridge (Cambridge, UK), 
Massachusetts Institute of Technology (Cambridge, MA USA), 
National Radio Astronomy Observatory (Charlottesville, VA USA), 
University of Pennsylvania (Philadelphia, PA USA), 
Scuola Normale Superiore di Pisa (Pisa, Italy),
SKA-South Africa (Cape Town, South Africa), and
University of Washington (Seattle, WA USA).  
Additional Collaborators are at 
Harvard University (Cambridge, MA USA),  
University of KwaZulu Natal (Durban, South Africa), 
University of Western Cape (Cape Town, South Africa), 
Imperial College London  (London, UK) and 
California State Polytechnic University (Pomona, CA USA).

This paper has the following structure.  Section \ref{sec:science} will give a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.  Key science results include
measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra});
characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra});
improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}); and
directly imaging large scale reionization structure (\S\ref{sec:imaging}).
Additionally, beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr});
searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}); and
performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).

Section \ref{sec:eormeas} will present the techniques used to make the measurement.  Section \ref{sec:requirements} will present the high-level requirements and Section \ref{sec:design} will present the system description..   Section \ref{sec:status} will present a brief status and outline the deployement timetable and Section \ref{sec:conclusion} will conclude by summarizing and providing additional context.

>>>>

The Hydrogen Epoch of Reionization Array (HERA; http://reionization.org) is a staged experiment to use the redshifted 21\,cm line
of neutral hydrogen to characterize our Cosmic Dawn, from
the formation of the first stars and black holes $\sim0.1~$Gyr after the Big
Bang ($z\sim30$) through the full reionization of the intergalactic medium
(IGM) $\sim1~$Gyr later ($z\sim6$).  By directly observing the large scale
structure of the primordial IGM as it is heated and reionized, HERA complements
probes at other wavelengths, adding transformative capabilities for
understanding the astrophysics and fundamental cosmology of our early universe.
Taking advantage of a new understanding of bright foreground systematics, HERA's
purpose-built radio
interferometer is optimized to deliver high signal-to-noise measurements of
redshifted 21\,cm emission to detect and characterize the Epoch of Reionization (EOR).

The power that observations of highly redshifted hydrogen emission have for answering key science about our early universe
has motivated a resurgence of interest in low-frequency arrays --- most with the primary
objective of measuring the EOR.  These include the Precision Array Probing the Epoch of Reionization (PAPER; \citealt{parsons_et_al2010}), the Murchison Widefield Array (MWA; \citealt{tingay_et_al2013}), the LOw Frequency ARray (LOFAR; \citealt{2013A&A...556A...2V}), and the Long Wavelength Array (LWA; \citealt{2009IEEEP..97.1421E}), as well as systems for existing dish arrays like the Very Large Array (VLA; \citealt{2013AAS...22132804K}) and the Giant Metrewave Radio Telescope (GMRT; \citealt{paciga_et_al2011}).
These experiments all struggle with the
inherent challenge of simultaneously meeting stringent sensitivity requirements
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal.
HERA improves on its predecessors by bringing significantly more sensitivity to bear
on the angular and spectral scales where recent work (discussed in Section \ref{sec:delayapproach}) has indicated that
the power spectrum of the EoR may dominate over foregrounds.
% XXX ARP I thought this was too early in the paper for the stuff below.  Relocate?
%At the same time, HERA continues to develop new techniques to extend and improve its capabilities.
%Examples include improving foreground subtraction to expand the information content 
%in power spectral and imaging measurements and
%extending HERA's performance at lower frequencies to probe IGM heating in the pre-reionization epoch (\S\ref{sec:EoX}).

@@

The HERA experiment will comprise 
350 14-m parabolic dishes (320 in a dense core $+$ 30 outriggers) in the South African Karoo Radio
Astronomy Reserve (see Fig. \ref{fig:HERApictures}, top, for a rendering).
HERA's antenna element and its compact configuration are
optimized for robust power spectrum detection,
delivering the requisite collecting area inexpensively, while carefully controlling foreground systematics.
The design and construction of this instrument is supported by the HERA collaboration 
consisting of the following Partner institutions:  
Arizona State University (Tempe, AZ USA), 
Brown University (Providence, RI USA), 
University of California Berkeley (Berkeley, CA USA), 
University of California Los Angeles (Los Angeles, CA USA), 
University of Cambridge (Cambridge, UK), 
Massachusetts Institute of Technology (Cambridge, MA USA), 
National Radio Astronomy Observatory (Charlottesville, VA USA), 
University of Pennsylvania (Philadelphia, PA USA), 
Scuola Normale Superiore di Pisa (Pisa, Italy),
SKA-South Africa (Cape Town, South Africa), and
University of Washington (Seattle, WA USA).  
Additional Collaborators are at 
Harvard University (Cambridge, MA USA),  
University of KwaZulu Natal (Durban, South Africa), 
University of Western Cape (Cape Town, South Africa), 
Imperial College London  (London, UK) and 
California State Polytechnic University (Pomona, CA USA).
The South African National Research Foundation SKA-SA group (SKA-SA) is a key partner in HERA's construction and science.

HERA's first stage of development has been funded under the US National Science Foundation's Mid-Scale Innovations Program,
which has supported the construction of a 19-element array for testing instrumental performance
and manufaturability on location.  
The first 19-element array is complete and, with additional funding from Cambridge University, 
construction is now underway on the next 18 elements.
The total 37 elements, to be completed in 2016, will provide a factor of about 5 more 
sensitivity than PAPER-128 and provide a significant chance at detecting the EoR power spectrum signal.  The next proposed phase will build out to 128 elements and use the existing 128 dual-polarization analog and digital signal paths that have been in use with PAPER,
%\citep{ali_et_al2015}
HERA-128 should provide a robust detection of the EOR signal and allow some characterization.  Finally, extending to 2020, HERA will build out to 350 elements to further EOR science as a function 
of redshift and spatial scale, potentially producing the first images of the EOR.  
%These estimates will be quantified and shown in this paper.

This paper has the following structure.  Section \ref{sec:science} gives a high-level overview of the science motivating HERA's construction, including primary and secondary science objectives.  Key science results include
measuring the 21\,cm power spectrum with high significance (\S\ref{sec:EoRPowerSpectra});
characterizing parametrized models of reionization and the first ionizing sources (\S\ref{sec:EoRPowerSpectra});
improving cosmic microwave background (CMB) constraints on primordial density fluctuations and neutrino masses by removing the optical depth degeneracy (\S\ref{sec:tau}); and
directly imaging large scale reionization structure (\S\ref{sec:imaging}).
Beyond these directly supported applications, HERA's public data products and instrumentation platform enable other high-impact science, including 
21\,cm cross-correlation (e.g. with JWST, WFIRST, CO, CII, and Ly$-\alpha$; \S\ref{sec:crossCorr});
searching for auroral radio bursts from exoplanetary magnetospheres (\S\ref{sec:exoplanets}); and
performing triggered, low-frequency follow up and characterization of fast radio bursts (\S\ref{sec:FRBs}).

Section \ref{sec:eormeas} presents the techniques used to make the measurement.  Section \ref{sec:requirements} presents the high-level requirements and Section \ref{sec:design} presents the system description..   Section \ref{sec:status} presents a brief status and outline the deployement timetable and Section \ref{sec:conclusion} concludes by summarizing and providing additional context.
